\chapter{Acknowledgments}

I would like to take this opportunity to express my sincere gratitude to all those who have contributed to the completion of this thesis. This journey has been both challenging and rewarding, and I am deeply appreciative of the support and guidance I have received along the way.

First and foremost, I extend my heartfelt thanks to my supervisor, Jasper Hoffman, for his unwavering support, mentorship, and invaluable insights throughout the research process. His dedication to pushing the boundaries of knowledge and commitment to excellence have been truly inspiring.

I would also like to acknowledge the significant contributions of my co-workers, Jan Ole von Hartz and Jens Rahnfeld, who helped also a lot in the development and debugging of the Variational Autoencoder (VAE) models used in this research. Their expertise, tireless efforts, and collaborative spirit greatly enriched the quality of this work.

I am grateful to my colleagues, friends, and family for their encouragement, patience, and understanding during this demanding endeavor. Your unwavering support provided the motivation needed to overcome challenges and persist in pursuit of our research goals.

Additionally, I extend my appreciation to the academic and research communities whose work has paved the way for this research. The wealth of knowledge and resources available in these communities has been instrumental in shaping the ideas presented in this thesis.

Lastly, I would like to acknowledge the resources provided by Neurobotics Lab lead by Professor Boedecker, which facilitated the successful execution of this research.

In closing, this research represents the collective efforts of many individuals who have contributed in various ways. Your support, expertise, and encouragement have been invaluable, and for that, I am truly grateful.